Parameter estimation

### 2022-05-27 SFE-report/Paramter_fitting.tex
27.05

% ===================================================
% METADATA
% ===================================================
\title[mode=title]{Parameter estimation}                      
\shorttitle{Forward sensitivity analysis of the supercritical extraction process}

@@

\newpage
\section{Parameter estimation} \label{CH: Parameter_estimation}
Estimation theory is a branch of statistics that deals with estimating the values of parameters based on measured empirical data that has a random component. The parameters describe an underlying physical setting in such a way that their value affects the distribution of the measured data. An estimator attempts to approximate the unknown parameters using the measurements.

For a given model, several statistical "ingredients" are needed so the estimator can be implemented. The first is a statistical sample – a set of data points taken from a random vector (RV) of size $N$. Let denote a statistical sample as $x$ and parameters whose vales are to be estimated as $\theta$. Both can be written into a vector form:
\begin{align}
	x &= [x[0], x[1], \dots, x[N-1] ]^T, &&x \in \mathbb{R}^N \\
	\theta &= [\theta_0, \theta_1, \dots, \theta_M ]^T,  &&\theta \in \mathbb{R}^M
\end{align}

where $N$ is the number of samples and $M$ is the number of parameter to estimation.

The continuous probability density function (pdf) or its discrete counterpart, the probability mass function (pmf), of the underlying distribution that generated the data must be stated conditional on the values of the parameters: $p(x|\theta)$. It is also possible for the parameters themselves to have a probability distribution (e.g., Bayesian statistics). It is then necessary to define the Bayesian probability $\pi(\theta)$. After the model is formed, the goal is to estimate the parameters, with the estimates commonly denoted $\hat{\theta}$.

One common estimator is the minimum mean squared error (MMSE) estimator, which utilizes the error between the estimated parameters and the actual value of the parameters as the basis for optimality. This error term is then squared and the expected value of this squared value is minimized for the MMSE estimator.

\subsection{Bayes theorem} \label{CH: Bayes_theorem}
In probability theory and statistics, Bayes' theorem (alternatively Bayes' law or Bayes' rule; recently Bayes–Price theorem[1]: 44, 45, 46 and 67 ), named after Thomas Bayes, describes the probability of an event, based on prior knowledge of conditions that might be related to the event.[2] For example, if the risk of developing health problems is known to increase with age, Bayes' theorem allows the risk to an individual of a known age to be assessed more accurately (by conditioning it on their age) than simply assuming that the individual is typical of the population as a whole.

One of the many applications of Bayes' theorem is Bayesian inference, a particular approach to statistical inference. When applied, the probabilities involved in the theorem may have different probability interpretations. With Bayesian probability interpretation, the theorem expresses how a degree of belief, expressed as a probability, should rationally change to account for the availability of related evidence. Bayesian inference is fundamental to Bayesian statistics.

\subsection{Maximum likelihood estimation} \label{CH: MLE}
\subsection{Maximum a posteriori} \label{CH: MAP}
\subsection{Bayesian inference} \label{CH: Bayesian_inference}

>>>>

% ===================================================
% METADATA
% ===================================================
\title[mode=title]{Parameter estimation}                      
\shorttitle{Parameter estimation}

@@

\newpage
\section{Bayes theorem} \label{CH: Bayes}

As discussed by \citet{Himmelblau1970}, The Bayesian approach to estimation make use of prior information. Such prior knowledge can come from theoretical considerations, from the results of previous experiments, or from assumptions by the experimenter. Typically, a Bayesian approach assumes a prior probability distribution of an unknown parameter $\theta$ in some parameter space $\boldsymbol{\theta}$. The distribution is updated by using Bayes' rule to obtain the posterior probability distribution.

Consider a set of events or outcomes, $A_1,A_2,...,A_n$, and some other event $B$. Bayes' theorem states that the probability that event $A_i$ will occur, given that event $B$ has already occured, which will denoted by $P\{A_i|B\}$, is equal to the product of the probability that $A_i$ will occur regardless of whether $B$ will take place and the probability that $B$ will occur, given that $A_i$  has already taken place, divided by the probability of the occurrence of $B$:

\begin{equation}
	P\{A_i|B\} = \cfrac{P\{B|A_i\}P\{A_i\}}{P\{B\}}
\end{equation}

Further, if al events comprising the set $\{A_i\}$ are included in $A_1,A_2,...,A_n$, then

\begin{equation} \label{EQ: Bayes_discrit}
	P\{A_i|B\} = \cfrac{P\{B|A_i\}P\{A_i\}}{ \sum_{i=1}^{n} P\{B|A_i\} P\{A_i\} }
\end{equation}

We can interpret these symbols as follows:

\begin{enumerate}
	\item $P\{A_i\}$ is a measure of our degree of belief that event $A_i$ will occur or that hypothesis $A_i$ is true prior to the acquisition of additional evidence that may alter the measure. $P\{A_i\}$ is denoted the \textit{prior probability}.
	\item $P\{A_i|B\}$ is a measure of our degree of belief that event $A_i$ will occur or that hypothesis $A_i$ is true, given additional evidence $B$ pertinent ot the hypothesis. $P\{A_i|B\}$ is termed the \textit{posterior probability}.
	\item $P\{B|A_i\}$ denotes the likelihood that event $B$ will occur, given that event $A_i$ is true. $P\{B|A_i\}$ is a conditional probability, interpreted in the Bayesian framework as a likelihood, $L(A_i|B)$.
\end{enumerate}

For continuous variable, Bayes' theorem can be more conveniently expressed in terms of the probability density function rather than the probabilities themselves. Equation \ref{EQ: Bayes_discrit} can be expressed in terms of a set of observed values of the random variable $X$, \textbf{x} and unknown parameter $(s)\theta$ as

\begin{equation} \label{EQ: Bayes_continous}
	p\left( \theta|X=\textbf{x} \right) = p\left( \theta|\textbf{x} \right) = \cfrac{L\left( \theta|\textbf{x} \right) p(\theta)}{\int_{-\infty}^{+\infty} L\left( \theta|\textbf{x} \right) p(\theta) d\theta}
\end{equation}

where

$p(\theta|\textbf{x}) = ~$the posterior probability density function for $\theta$; it includes knowledge of the possible values of $\theta$ gained from the experimental data \textbf{x}

$p(\theta) = ~$ the prior probability density function for $\theta$ (before the experiment in which \textbf{x} was observed)

$L\left( \theta|\textbf{x} \right) = p\left( \textbf{x}|\theta \right) = ~$ the probability density function termed the likelihood function of $\theta$ given \textbf{x}

The denominator in Equation \ref{EQ: Bayes_continous} is a normalizing factor chosen so that the integration of the posterior distribution is unit, i.e., $\int_{-\infty}^{+\infty} p\left( \theta|\textbf{x} \right) = 1$. By taking into account the law of total probability, the denominator can be written as

\begin{equation}
	\int_{-\infty}^{+\infty} p\left( \textbf{x}|\theta \right) p(\theta) d\theta = p(\textbf{x})
\end{equation}

If the prior distribution is a uniform distribution, that is the prior distribution is a constant, the \ref{EQ: Bayes_continous} reduces to

\begin{equation}
	p\left( \theta|\textbf{x} \right) = \cfrac{L\left( \theta|\textbf{x} \right)}{\int_{-\infty}^{+\infty} L\left( \theta|\textbf{x}\right) d\theta}
\end{equation}

If the prior knowledge concerning a postulated event or hypothesis is poor, the posterior probability is largely or entirely determined by the likelihood function, that is, by the additional accumulated evidence for which the likelihood function acts as a mathematical expression. If prior knowledge outweighs recent evidence, however, then the posterior probability is determined almost solely by the prior probability.

In the application of tests and the design of experiments, certain definitions and rules concerning probability are needed and are listed below.

\begin{enumerate}
	\item It follows from the frequency theory of probability that: $0 \leq P \leq 1$
	\item If the probability of occurrence of one event $A$ depends on whether or not event $B$ has occurred, the two events are termed \textit{dependent}; if the probability of occurrence of event $A$ does not depend on the occurrence of $B$, or the reverse, the two events are \textit{independent}.
	\item \textbf{Addition Rule}
	If $A_1,A_2,...,A_n$ are mutually exclusive events, i.e., cannot occur at the same time, the probability of occurrence of just one of the events is equal to the sum of the probabilities of each $A_i$:

	\begin{equation}
		P\left( A_1, \text{or~} A_2,..., \text{or~} A_n \right) = \sum_{i=1}^{n} P\left( A_i \right)
	\end{equation}

	Very often we let
	\begin{equation}
		\sum_{i=1}^{n} P\left( A_i \right) = 1
	\end{equation}

	Also, if each event is equi-probable so that $P(A_i) = q$,

	\begin{equation}
		\sum_{i=1}^{n} q = nq = 1 \qquad \text{or} \qquad  q=\cfrac{1}{n} = P\left( A_i \right)
	\end{equation}

	In set theory, mutually exclusive event have no points in common. The union of the sets which represents the set of all elements that belong to

	\begin{equation}
		P\left( A_1 \cup A_2 \cup ... \cup A_n \right) = P(A_1) + P(A_2) + ... + P(A_n)
	\end{equation}

	\item \textbf{Multiplication Rule}
	If $A$ and $B$ are \textit{independent} event

	\begin{equation}
		P\left( A \text{~and~} B\right) = P(A)P(B)
	\end{equation}

	In set theory the intersection of $A$ and $B$ is the set of all elements that belong to $A$ and $B$:

	\begin{equation} \label{EQ: Probabiltiy_independet_a}
		P\left( A \cap B  \right) = P(A)P(B)
	\end{equation}

	If the $A$ and $B$ are \textit{dependet} events,

	\begin{equation} \label{EQ: Probabiltiy_independet_b}
		P\left(A|B\right) = \cfrac{P\left( A \cap B \right)}{P(B)}
	\end{equation}

	where the symbol $P(A|B)$ means "probability of $A$ given $B$". As a corollary,

	 \begin{subequations}
		\begin{alignat}{2}
			P\left( A \cap B \right) &= P(B) P(A|B) \label{EQ: Probabiltiy_dependet_a} \\
			&= P(A)P(B|A) \label{EQ: Probabiltiy_dependet_b}
		\end{alignat}
	\end{subequations}

	Two kind of probabilities enter Equation \ref{EQ: Probabiltiy_dependet_a} (or Equation \ref{EQ: Probabiltiy_dependet_b}): the absolute probability of event $B$ (or $A$) irrespective of whether or not $A$ (or $B$) has occurred, and that the conditional probability of event $A$ (or $B$) computed on the assumption that $B$ (or $B$) has occurred. Equation \ref{EQ: Probabiltiy_independet_a} or \ref{EQ: Probabiltiy_independet_b} is special case of Equation \ref{EQ: Probabiltiy_dependet_a} or \ref{EQ: Probabiltiy_dependet_b}, because if the event are independent $P(A|B)=P(A)$.
	For the case of many event, Equation \ref{EQ: Probabiltiy_independet_a} can be expanded to

	\begin{equation}
		\begin{split}
			P(A_1 \text{~and~} A_2 \text{~and~} ... \text{~and~} A_n) &= P(A_1)\cdot P(A_2) \cdot  ...\cdot P(A_n) \\ &= \prod_{i=1}^{n} P(A_i)
		\end{split}
	\end{equation}

	\item Another useful relationship for event which are not mutually exclusive is

	\begin{equation}
		P(A) + P(B) - P(A \cap B) = P(A \cup B)
	\end{equation}

\end{enumerate}

\newpage
\section{Parameter estimation} \label{CH: Parameter_estimation}

Conceptually, in the measurement equation the unobservable $\epsilon(t)$ is added to $y(t)$ to give the observable dependent variable $Y(t)$ defined as:

\begin{equation}
	Y(t) = y(t) + \epsilon(t)
\end{equation}

If the estimated parameter $\hat{\alpha}$ replaces the model parameter $\alpha$, the residue error is $\textbf{E} = Y(t) - \hat{Y}(t)$. The objective in parameter estimation is to obtain the "best" estimate of $\alpha$ based on the continuous observations $Y(t)$ or the discrete observations $Y(t_i)$.

We assume that $x(t)$ is given \textit{a priori} but the $\textbf{y}_0$ and $\alpha$ are to be estimated over the time interval $0 \leq t \leq t_n$ from discrete observation described by the following relation:

\begin{equation}
	\textbf{Y}(t_i) = h(t_i)y(t_i) + \epsilon(t_i) \qquad 1 \leq i \leq n
\end{equation}

where $\textbf{Y}(t_i)$ is an $n \times 1$ column vector, $h(t_i)$ is an $n \times v$ matrix given \textit{a priori}, and $\epsilon (t_i)$ is an $x \times 1$ column vector ("noise" vector) whose elements are the unobservable errors. The above equation can be written in general form as:

\begin{equation} \label{EQ: Measurment_noise}
	\textbf{Y}(t_i) = \Psi(\alpha, y_0, t_i) + \epsilon(t_i)
\end{equation}
where $\Psi$ is the matrix of functions.

Because of the difficulty of obtaining analytical solutions to the deterministic process model, experiments have been arranged wherby the vector of derivatives $d\textbf{Y}/dy$ is measured rather than \textbf{Y} itself. In such cases, it is assumed that the unobservable error is added to the deterministic derivative $dy/dt$ as follows:

\begin{equation}
	\cfrac{d\textbf{Y}}{dt} = \cfrac{dy}{dt} + \epsilon
\end{equation}

\subsection{Least squares estimation}
A least squares parameter estimation does not require prior knowledge of the distribution of unobservable errors, yield  unbiased estimates, that is, $ \mathbb{E}\{\hat{Y}(t) = y(t) \} $, and results in the minimum variance among all linear unbiased estimators. If the observations $\textbf{Y}$ for the model responses are continuous functions of time from $t = 0$ to $t = t_f$, the Markov (or "rigorous least squares") criterion is to minimize:

\begin{equation}
	\phi =  \cfrac{1}{2} \int_{0}^{t_f}  \left[ \textbf{Y} - \Psi \right]^T \Gamma^{-1} \left[ \textbf{Y} - \Psi \right] dt
\end{equation}

where $\Gamma$ is the covariance matrix (or perhaps a matrix of appropriate weights), and $\phi$ is the time integrated value of the error squared ("integral squared error"). If the observations are made at discrete instants of time, $t_i$, $i = 1,2,...,n$, the Markov criterion is to minimize:

\begin{equation}
	\phi = \cfrac{1}{2} \sum_{i=1}^{n} \left[ \textbf{Y}(t_i) - \Psi(t_i) \right]^T \Gamma^{-1} \left[ \textbf{Y}(t_i) - \Psi(t_i) \right] 
\end{equation}

If $\Gamma$ is a diagonal matrix, $\phi$  becomes a "weighted least squares" criterion; if $\Gamma = \sigma^2 \textbf{I}$ where $\sigma$ is a standard deviation, $\phi$ is the "ordinary least squares" criterion.

\subsection{Maximum likelihood estimation}

Maximum likelihood estimation (MLE) is a method of estimating the parameters of an assumed probability distribution, given some observed data. This is achieved by maximizing a likelihood function so that, under the assumed statistical model, the observed data is most probable. The MLE has the desirable characteristics of asymptotic efficiency and normality. Each time it has been associated with the (joint) normal distribution because of mathematical convenience. Consider the joint probability density function (the likelihood function) $p(\alpha, y_0|y(t_1),y(t_2),...,y(t_n))$ for $\alpha$ and $y_0$ if a maximum of this function over all choices of $y_0$ and $\alpha$ can be found, the estimates so obtained are maximum likelihood estimates. The conditions at the maximum can be evolved incorporating prior information as follows.

The posterior probability density $p(\alpha,y_0|y(1),y(2),...,y(t_n))$ can be expressed as the ratio of two probability densities if we make use of the analog for continuous variables of Equation \ref{EQ: Probabiltiy_independet_b}:

\begin{equation} \label{EQ: Posterior_distribution}
	p\left( \alpha,y_0|y(t_n,...,y(1)) \right) = \cfrac{p\left( \alpha, y_0, y(t_n),...,y(t_1) \right)}{p\left( y(t_n),...,y(t_1) \right)}
\end{equation}

The numerator of the right-hand side od Equation \ref{EQ: Posterior_distribution} using Equation \ref{EQ: Probabiltiy_dependet_a} becomes

\begin{equation}
	\begin{split}
		p\left( \alpha,y_0,y(t_n),...,y(t_1) \right) &= p\left( y(t_n)|\alpha,y0,y(t_{i-1},...,y(t_1)) \right) \\ &\cdot p\left( \alpha,y_0,y(t_{n-1}),...,y(t_1) \right)
	\end{split}
\end{equation}

These operations can be continued repetitively until we get

\begin{equation}
	p\left( \alpha,y_0,y(t_n),...,y(t_1) \right) = p\left( \alpha,y_0 \right) \prod_{i=1}^{n} p\left( \alpha,y_0,y(t_{n-1}),...,y(t_1) \right)
\end{equation}

Examination of Equation \ref{EQ: Measurment_noise} shows that $\textbf{Y}(t_i)$ depends only on $t_i$, $y_0$, $\alpha$ and $\epsilon(t_i)$ and is not conditioned by any previous measurement. Consequently, we can write

\begin{equation}
	p\left( y(t_i)|\alpha,y_0, y(t_{i-1}),...,y(t_1) \right) = p\left( y(t_i)|\alpha,y_0 \right)
\end{equation}

provided Equation \ref{EQ: Measurment_noise} is observed as a constraint. The desired joint conditional probability function is thus

\begin{equation}
	p\left( \alpha,y_0|y(t_n),...,y(t_1) \right) = \cfrac{p\left( \alpha,y_0 \right) \prod_{i=1}^{n} p\left( y(t_i)|\alpha,y_0 \right) }{p\left( y(t_n),...,y(t_1) \right)}
\end{equation}

The function to be maximized will be the logarithm of the likelihood function $L = p\left( \alpha,y_0, y(t_n),...,y(t_1) \right)$ constrained by equation \ref{EQ: Measurment_noise}, which can be written as $\ln(L)$ plus the Lagrangian multipliers $\lambda(t_k)$ times the constraint function, or

\begin{equation} \label{EQ: MLE_Lagrangian}
	\begin{split}
		L^* \equiv \ln p\left( \alpha,y_0 \right) + \sum_{i=1}^{n} \left\{ \ln p(y(t_i)|\alpha,y_0) \right. \\
		\left. + \lambda^T \cdot \left[ y(t_i) - \Psi\left( \alpha,y_0,t_i \right) - \epsilon(t_i) \right] \right\} \\ - \ln \left[ p\left( y(t_n),...,y(t_1) \right) \right]
	\end{split}
\end{equation}

By assumption of the relation of Equation \ref{EQ: Measurment_noise}

\begin{equation} \label{EQ: Noise_prob}
	p\left( y(t_i) | \alpha,y_0 \right) = p\left( \epsilon(t_i) \right)
\end{equation}

and specially

\begin{equation} \label{EQ: Noise_prob_norm}
	p\left( \epsilon(t_i) \right) = \cfrac{1}{ (2\pi)^{v/2} |\Gamma(t_i)|^{1/2} } \exp \left[ -\cfrac{1}{2}\epsilon(t_i) \left( \Gamma(t_i) \right)^{-1} \epsilon(t_i) \right]
\end{equation}
where $|\Gamma|$ is the det$~\Gamma$, and $\Gamma$ is the covariance matrix of $\epsilon$, i.e., of the response.

After Equation \ref{EQ: Noise_prob} is substituted into Equation \ref{EQ: MLE_Lagrangian}, and $L^*$ is differentiated with respect to each of the estimated and $\epsilon(t_i)$, and the resulting expression is equated to zero, we get

 \begin{subequations} \label{EQ: MLE_Lag_diff}
	\begin{alignat}{2}
		\cfrac{\partial L^*}{\partial y_0} &= \cfrac{\partial}{\partial y_0} \ln p(\alpha,y_0) - \sum_{i=1}^{n} \lambda^T(t_i) \cfrac{\partial}{\partial y_0} \Psi(\alpha, y_0,t_i) = 0^T \\
		\cfrac{\partial L^*}{\partial \alpha} &= \cfrac{\partial}{\partial \alpha} \ln p(\alpha,y_0) - \sum_{i=1}^{n} \lambda^T(t_i) \cfrac{\partial}{\partial \alpha} \Psi(\alpha, y_0,t_i) =0 \\
		\cfrac{\partial L^*}{\partial \epsilon(t_i)} &= \cfrac{\partial}{\partial \epsilon(t_i)} \ln p(\epsilon(t_i)) - \lambda^T(t_i) = 0^T
	\end{alignat}
 \end{subequations}

Substitution of Equation \ref{EQ: Noise_prob_norm} into the last Equation of \ref{EQ: MLE_Lag_diff} makes it possible to solve for $\lambda(t_i)$:

\begin{align}
	\lambda(t_i) &= -\left( \Gamma(t_i) \right)^{-1} \epsilon(t_i) \nonumber \\
	&= -\left( \Gamma(t_i) \right)^{-1} \left[ \textbf{Y}(t_i) - \Psi(\alpha,y_0,t_i) \right]
\end{align}

and to eliminate $\lambda(t_i)$ from the first two equations of \ref{EQ: MLE_Lag_diff}.

For convenience we shall define a new column vector $\alpha^*$ in which all the elements of $\alpha$ are arranged as follows:

\begin{equation}
	\alpha^* = \begin{bmatrix}
		\alpha_{1,1} & \alpha_{1,2} & \cdots & \alpha_{1,v} & \alpha_{2,1} & \cdots & \alpha_{2,v} & \cdots & \alpha_{v,v}
	\end{bmatrix}^T	
\end{equation}

Finally, we assume that $y_0$ and $\alpha^*$ are distributed by a joint normal distribution and that the prior  distribution are, respectively

\begin{subequations} \label{EQ: MLE_Priors}
	\begin{align}
		&p\left( \alpha^* \right) = \cfrac{1}{ (2\pi)^{v/2} |\Omega_{\alpha^*}|^{1/2} } \exp \left[ -\cfrac{1}{2} \left( \alpha^* - \alpha^{*(0)} \right)^T \Omega_{\alpha^*}^{-1} \left( \alpha^* - \alpha^{*(0)} \right) \right] \\
		&p\left( y_0 \right) = \cfrac{1}{ (2\pi)^{v/2} |\Omega_{y_0}|^{1/2} } \exp \left[ -\cfrac{1}{2} \left( y_0 - y_0^{(0)} \right)^T \Omega_{y_0}^{-1} \left( y_0 - y_0^{(0)} \right) \right] 
	\end{align}
\end{subequations}
where $\Omega$'s are the respective covariance matrices for $\alpha^*$ and $y_0$, and the subscript $(0)$ designates the prior estimated of $\alpha^*$ and $y_0$. If we assume that $\alpha^*$ and $y_0$ are independent

\begin{equation}
	\ln p(\alpha^*,y_0) = \ln p(\alpha^*) + \ln p(y_0)
\end{equation}

Introduction of the prior distributions, Equations \ref{EQ: MLE_Priors}, plus the expression for $\lambda(t_i)$ into the first two equations of \ref{EQ: MLE_Lag_diff} gives the final equations from which the estimators of $\alpha^*$ and $y_0$ can be obtained:

\begin{subequations} 
	\begin{align}
	-\left( \hat{y}_0 - y_0^{(0)} \right)^T \Omega_{y_0}^{-1} &+ \sum_{i=1}^{n}\left[ \textbf{Y}(t_i) - \Psi\left( \hat{\alpha}, \hat{y}, t_i \right) \right]^T \Gamma^{-1}(t_i) \nonumber \\ 
	&\cdot \cfrac{\partial}{\partial y_0} \Psi\left( \hat{\alpha}, \hat{y}, t_i \right) = 0^T \\
	-\left( \hat{\alpha}^* - \alpha^{*(0)} \right)^T \Omega_{\alpha^*}^{-1} &+ \sum_{i=1}^{n}\left[ \textbf{Y}(t_i) - \Psi\left( \hat{\alpha}, \hat{y}, t_i \right) \right]^T \Gamma^{-1}(t_i) \nonumber \\ 
	&\cdot \cfrac{\partial}{\partial \alpha^*} \Psi\left( \hat{\alpha}, \hat{y}, t_i \right) = 0^T \nonumber \\
	\end{align}
\end{subequations}
where the overlay caret denotes estimated parameters.

Note that under the assumption that the elements of $\Omega$ are essentially infinite (prior knowledge is diffuse), the elements of $\Omega^-1$ are zero, the equations for the maximum likelihood estimates coincide with those for the least squares estimates, and the same calculations for precision in the estimates apply.

\hrule

For some models, these equations can be explicitly solved for $\hat{\alpha}^*$ and $\hat{y}$ but in general no closed-form solution to the maximization problem is known or available, and an MLE can only be found via numerical optimization. Under the assumption of normality, the lack of prior knowledge about the parameters and known initial conditions, the optimization problem is defined as follow:

\begin{equation}
	\hat{\alpha} = \arg \max_{\alpha \in \Theta} L = \arg \max_{\alpha \in \Theta} p(\alpha|y(t_1),y(t_2),...,y(t_n))
\end{equation}

Intuitively, this selects the parameter values that make the observed data most probable. The specific value $\hat{\alpha}$ that maximizes the likelihood function $L$ is called the maximum likelihood estimate. Further, if the function $\hat{\alpha}:\mathbb{R}\rightarrow\Theta$ so defined is measurable, then it is called the maximum likelihood estimator.

If the random samples $y(t_1)$, $y(t_2)$,...,$y(t_n)$ come from a normal distribution with unknown mean $\hat{\alpha}$ and variance $\sigma^2$, then the joint probability function can be written as

\begin{equation}
	p(\alpha|y(t_i)) = \cfrac{1}{\sqrt{2\pi\sigma}}\exp\left[-\cfrac{ \left( \textbf{Y}(t_i) - \Psi\left( \hat{\alpha}, \hat{y}, t_i \right) \right)^2}{2\sigma}\right]
\end{equation}

Then the likelihood function is defined as follow:

\begin{multline} 
	L =\prod\limits_{i=1}^n p(\alpha|y(t_i))= \\ \sigma^{-n/2}(2\pi)^{-n/2}\exp\left[-\cfrac{1}{2\sigma}\sum\limits_{i=1}^n \left( \textbf{Y}(t_i) - \Psi\left( \hat{\alpha}, \hat{y}, t_i \right) \right)^2\right]
\end{multline}

By taking the ln of the likelihood function, the final form of a cost function is obtained:

\begin{equation} 
	\ln L=-\cfrac{n}{2}\ln\sigma-\cfrac{n}{2}\ln(2\pi)-\cfrac{\sum\limits_{i=1}^n \left( \textbf{Y}(t_i) - \Psi\left( \hat{\alpha}, \hat{y}, t_i \right) \right)^2}{2\sigma}
\end{equation}
